### 2015-05-03 draft.tex
..

% Including an example policy, todo change 0 to 10 to actual IP
Figure~\ref{fig:firewall-eg} displays a policy for a stateful firewall using our API. Sources from 0 to 10 are taken to be internal to the network. If a packet is sent from an internal host to an external host, the pair of internal and external host is added to a list of trusted connection. If a packet is received from an external host, we lookup the table to see if the connection is trusted, in which case we forward the packet to its destination, otherwise we drop it.

\begin{figure}
\begin{lstlisting}
f(pkt):
  if Oak.in_range(pkt, src, 0, 10) then
     if Oak.in_range pkt dst  [(0,10)] then
        Oak.deliver
     else 
        Oak.add pkt (Oak.IpDst, Oak.IpSrc) ``conn'';
        Oak.deliver
  else if Oak.in_rel pkt (Oak.IpSrc,Oak.IpDst) ``conn'' then
         Oak.deliver
       else
         RO.drop
       end
  end                
  \end{lstlisting}
\label{fig:firewallcode}
\caption{A Stateful Firewall}
\end{figure}

@@

  \begin{figure}
\label{fig:decisiontree}  
\caption{Decision Tree for Stateful Firewall}     
\end{figure}

@@

   \begin{figure}
     \begin{lstlisting}

function InRange: pkt * field * value * value -> bool

\end{lstlisting}

\caption{Signature of the relation portion of the API}
\end{figure}

@@

\subsection*{A.P.I.}
In our model, algorithmic policies interact with packets through A.P.I function calls before returning a forwarding decision. We include in Figure~\ref{fig:range_api} the core of our \lstinline|API|.

\lstinline|field| enumerates the different fields of a packet. It can be extended to represent more fields of a packet's header. Associated with \lstinline|field| are \lstinline|range|s, possible values taken by a certain packet's field, within a given minimum value \lstinline|min_value| and maximum value \lstinline|max_value|. \lstinline|range| consists of list of integer pairs, representing a collection of intervals whose union correspond to the range of values of a packet. For simplicity, we consider integer fields here, but this could be extended, for example, to IP prefix, boolean flags, etc.

@@

\begin{figure}
  \begin{lstlisting}
(* abstract packet type *)
type packet

@@

(* in_range p f r returns true if the value of field f of packet p is in range r, and false otherwise *)
val in_range: packet -> field -> range -> bool
\end{lstlisting}
\label{fig:range_api}
\caption{Signature of the packet and range-branching portion of the API}
\end{figure}

\begin{figure}
  \begin{lstlisting}
(* name of a relation *)
type relation = string

@@

(* test if a packet's fields are in a relation *)
val in_relation: packet -> field list -> relation -> bool
\end{lstlisting}
\label{fig:rel_api}
\caption{Signature of the relation portion of the API}
\end{figure}

\begin{figure}
  \begin{lstlisting}
(* compile a user-function into a policy *)
val compile: (packet -> forwarding_decision) -> policy

@@

(* Print the policy to stdout *)
val print_policy: policy -> unit
  \end{lstlisting}
  \label{fig:build_api}
  \caption{Signature of the interactive portion of the API}
\end{figure}

@@

\begin{figure}
\begin{lstlisting}
let compile (f: packet -> forwarding_decision) : policy =
      push most general packet to (empty) stack
      set the current location to be the root of an empty decision tree
      while (stack is not empty)
         pop the top of the stack as pkt
         run f of pkt resulting in a forwarding decision fd (and a modified packet pkt)
         record the forwarding decision for the symbolic packet pkt at the current location in the decision tree
         set the current location to be the root of the decision tree and repeat

@@

When \lstinline|f| calls \lstinline|in_range| to learn if the field \lstinline|fld|'s value is in the range \lstinline|r|, we (1) create an \lstinline|Inrange| branching node in the decision tree (2) refine the current symbolic packet to represent all packets in range \lstinline|r| that could reach this test in the algorithmic policy and return true (3) push on the stack of packets to process a symbolic packets representing all packets in the complement of \lstinline|r| for field \lstinline|fld|.

Similarly, when the algorithmic policy calls \lstinline|in_relation| to learn if fields \lstinline|flds|'s value are in a relation \lstinline|rel|, we (1) create an \lstinline|Inrelation| branching node in the decision tree (2) tag the current packet as being in the relation, return true (3) push on the stack of packets to process the same packet tagged as not being part of relation \lstinline|rel|.

When \lstinline|f| calls \lstinline|add| or \lstinline|remove|, we record in the decision the current symbolic packet, the relation and the fields being added or removed from the relation before continuing with the execution of \lstinline|f|.

@@

\begin{figure}
\begin{lstlisting}
let run (pol: policy) (inputs: ((field*int) list) list) : unit =
      build the symbolic rules for policy pol
      generate the concrete rules with all relations empty
      for each packets in input,
          find the first of the current rule matching the packet, and record the forwarding decision fd
          if packet is sent to the controller then
              evaluate the policy on the packet, potentially adding or removing tuples from relations, record forwarding decision fd'
              regenerate concrete rules
              forward the packet according the forwarding decision fd'
          else
              forward the packet according to fd 
\end{lstlisting}

>>>>

Figure~\ref{fig:firewall-eg} displays a policy for a stateful firewall using our API. Sources from 0 to 10 are taken to be internal to the network. If a packet is sent from an internal host to an external host, the pair of internal and external host is added to a list of trusted connection. If a packet is received from an external host, we lookup the table to see if the connection is trusted, in which case we forward the packet to its destination, otherwise we drop it.

\begin{figure}[ht]
\begin{lstlisting}
f(pkt):
  if Oak.in_range(pkt, src, 0, 10) then
     if Oak.in_range pkt dst  [(0,10)] then
        Oak.deliver
     else 
        Oak.add pkt (Oak.IpDst, Oak.IpSrc) ``conn'';
        Oak.deliver
  else if Oak.in_rel pkt (Oak.IpSrc,Oak.IpDst) ``conn'' then
         Oak.deliver
       else
         RO.drop
       end
  end                
  \end{lstlisting}

\caption{A Stateful Firewall}
\label{fig:firewallcode}
\end{figure}

@@

  \begin{figure}

\caption{Decision Tree for Stateful Firewall}     
\label{fig:decisiontree}  
  \end{figure}

@@

@@

\subsection*{A.P.I.}
In our model, algorithmic policies interact with packets through A.P.I function calls before returning a forwarding decision. We include in Figure~\ref{fig:range_api} the core of our \lstinline|A.P.I.|. 
\lstinline|field| enumerates the different fields of a packet. It can be extended to represent more fields of a packet's header. Associated with \lstinline|field| are \lstinline|range|s, possible values taken by a certain packet's field, within a given minimum value \lstinline|min_value| and maximum value \lstinline|max_value|. \lstinline|range| consists of list of integer pairs, representing a collection of intervals whose union correspond to the range of values of a packet. For simplicity, we consider integer fields here, but this could be extended, for example, to IP prefix, boolean flags, etc.

@@

Algorithmic policy may base their forwarding decision on the result of calling the \lstinline|in_range pkt fld r|, which returns whether value of packet \lstinline|pkt| at field \lstinline|fld| is in the range \lstinline|r|.

\begin{figure}[ht]
  \begin{lstlisting}
(* abstract packet type *)
type packet

@@

(* in_range p f r returns true if the value of field f of packet p is in range r, and false otherwise *)
val in_range: packet -> field -> range -> bool
\end{lstlisting}

\caption{Signature of the packet and range-branching portion of the API}
\label{fig:range_api}
\end{figure}

In Figure~\ref{fig:rel_api}, we include the relation portion of the \lstinline|A.P.I.|, used to model stateful controller programs. A \lstinline|relation| represents a table in which each row contains a series of field values. It is referred to by a string. Policy writer may add a row to a relation \lstinline|rel| by calling the function \lstinline|add pkt flds rel|, which will populate the new row with the fields \lstinline|flds|' values of the current packet \lstinline|pkt|. Similarly, they may remove a row containing exactly the values of fields \lstinline|flds|' values of the current packet \lstinline|pkt| by calling \lstinline|remove pkt flds rel|. Finally, one may test if the values of fields \lstinline|flds| of current packet \lstinline|pkt| are in relation \lstinline|rel| by calling \lstinline|in_relation pkt flds rel|.

\begin{figure}[ht]
  \begin{lstlisting}
(* name of a relation *)
type relation = string

@@

(* test if a packet's fields are in a relation *)
val in_relation: packet -> field list -> relation -> bool
\end{lstlisting}

\caption{Signature of the relation portion of the API}
\label{fig:rel_api}
\end{figure}

\begin{figure}[ht]
  \begin{lstlisting}
(* compile a user-function into a policy *)
val compile: (packet -> forwarding_decision) -> policy

@@

(* Print the policy to stdout *)
val print_policy: policy -> unit
  \end{lstlisting}

  \caption{Signature of the interactive portion of the API}
  \label{fig:build_api}
\end{figure}

@@

\begin{figure}[ht]
\begin{lstlisting}
let compile (f: packet -> forwarding_decision) : policy =
      push most general packet to (empty) stack
      set the current location to be the root of an empty decision tree
      while (stack is not empty)
         pop the top of the stack as pkt
         run f of pkt resulting in a forwarding decision fd (and a modified packet pkt)
         record the forwarding decision for the symbolic packet pkt at the current location in the decision tree
         set the current location to be the root of the decision tree and repeat

@@

When \lstinline|f| calls \lstinline|in_range| to learn if the field \lstinline|fld|'s value is in the range \lstinline|r|, we
\begin{enumerate}
\item  create an \lstinline|Inrange| branching node in the decision tree
\item refine the current symbolic packet to represent all packets in range \lstinline|r| that could reach this test in the algorithmic policy and return true
\item push on the stack of packets to process a symbolic packets representing all packets in the complement of \lstinline|r| for field \lstinline|fld|.
\end{enumerate}

Similarly, when the algorithmic policy calls \lstinline|in_relation| to learn if fields \lstinline|flds|'s value are in a relation \lstinline|rel|, we
\begin{enumerate}
  \item create an \lstinline|Inrelation| branching node in the decision tree
  \item tag the current packet as being in the relation, return true
  \item push on the stack of packets to process the same packet tagged as not being part of relation \lstinline|rel|.
\end{enumerate}
When \lstinline|f| calls \lstinline|add| or \lstinline|remove|, we record in the decision the current symbolic packet, the relation and the fields being added or removed from the relation before continuing with the execution of \lstinline|f|.

@@

\begin{figure}[ht]
\begin{lstlisting}
let run (pol: policy) (inputs: ((field*int) list) list) : unit =
      build the symbolic rules for policy pol
      generate the concrete rules with all relations empty
      for each packets in input,
          find the first of the current rule matching the packet, and record the forwarding decision fd
          if packet is sent to the controller then
              evaluate the policy on the packet, potentially adding or removing tuples from relations, record forwarding decision fd'
              regenerate concrete rules
              forward the packet according the forwarding decision fd'
          else
              forward the packet according to fd 
\end{lstlisting}
